---
title: "Visualizing data on a map with {ggplot2} and {sf}"
author: "Johannes Enevoldsen"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "figs/",
  message = FALSE,
  warning = FALSE
)
```
  
  > This document generated is from `README.rmd`

This vignette demonstrates how to draw maps of different administrative divisions of Denmark (municipalities and regions), and how to decorate the map with geographically anchored data.

The shapefile used is *Danmarks Administrative Geografiske Inddeling 1:2 mio* downloaded from [Styrelsen for Dataforsyning og Effektivisering](kortforsyningen.dk). This data can be used and redistributed according to [terms and conditions](https://download.kortforsyningen.dk/content/vilk%C3%A5r-og-betingelser).

We will use the package [{sf}](https://r-spatial.github.io/sf/) to handle the geospatial data and [{ggplot2}](https://ggplot2.tidyverse.org/) to draw it.

```{r}
library(tidyverse) # loads ggplot2 and a number of other packages incl. dplyr and readr
library(sf)
```

First, load the municipality ("KOMMUNE") layer from the shapefile. Other relevant layers are "REGION" and "SOGN".

```{r}
sf_municipality <- read_sf("SHAPEFILE-DK/", layer = "KOMMUNE")
```

This returns a data frame (tibble) with some metadata. Here, we are mainly interested in the columns `KOMNAVN` (name of municipality) and `geometry`.

Let's draw.

```{r}
ggplot(sf_municipality) + 
    geom_sf() + # geom_sf uses the aesthetic geometry instead of x and y. 
    # By default it will assume it is stored in the column geometry.
    coord_sf()
```

Adding data is just as simple. To demonstrate, we will color each municipality according to the total number of contacts to a general practitioner (data from [Danmarks Statistik](dst.dk)).

```{r}
gp_contacts <- read_csv("gp-contacts-2020.csv", 
                        col_names = c("Kommune", "n_contacts"),
                        locale = locale(encoding = "latin1"))

knitr::kable(head(gp_contacts))
```

Now we join the data and the shapes. First, we should check that the same municipalities are present in both data frames (and that they are spelled the same way).

```{r}
c(
"sf - gp" = setdiff(sf_municipality$KOMNAVN, gp_contacts$Kommune),
"gp - sf" = setdiff(gp_contacts$Kommune, sf_municipality$KOMNAVN)
)
```

Oh well... [Christiansø](https://da.wikipedia.org/wiki/Christians%C3%B8) is a small island northeast of Bornholm. It has 90 inhabitants, it is not really a municipality, and not part of another municipality. It's administered by the Departement of Defense. We can probably safely ignore this.

Join the GP data to the sf data. 

  > ⚠️ The order in the join statement is important (sf data first), since the returned variable `gp_contacts_w_sf` must be of class "sf" for `geom_sf()` and `coord_sf()` to do its magic 🪄

```{r}
gp_contacts_w_sf <- left_join(sf_municipality, gp_contacts, by = c("KOMNAVN" = "Kommune"))

ggplot(gp_contacts_w_sf, aes(fill = n_contacts)) + 
    geom_sf(aes(fill = n_contacts)) +
    coord_sf() +
    scale_fill_viridis_c(label = scales::label_number_si()) + 
    labs(title = "Number of contacts to a general practitioner in 2020",
         fill = "# contacts")
```
