---
title: "Stringency Index - Oxford Covid-19 Government Response Tracker (OxCGRT)"
author: "Johannes Enevoldsen"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "figs/",
  message = FALSE,
  warning = FALSE,
  print
)
```
  
  > This document is generated from [README.rmd](README.rmd)
  
The OxCGRT stringency index is an attempt to quantify the degree to which a given government enforces restrictions on its citizens with the aim of reducing the spread of COVID-19.

From the [official website](https://covidtracker.bsg.ox.ac.uk/):
  
  > Governments are taking a wide range of measures to tackle the COVID-19 outbreak. We aim to track and compare worldwide government responses to the coronavirus rigorously and consistently. Systematic information on which measures governments take, and when, can help us understand the responses in a consistent way, aiding efforts to fight the pandemic.
  
  
The [official website](https://covidtracker.bsg.ox.ac.uk/) provide a lot of information about the different types of restrictions that go into the stringency index. Timeseries about the individual restrictions can also be accessed.

Here we will download an plot the timeseries of the stringency index.

```{r}
library(tidyverse)

stringency_index_url <- "https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/timeseries/stringency_index.csv"

stringency_index <- read_csv(stringency_index_url)

print(stringency_index, n_extra = 10)
```

The data is in a wide format, with dates in column headers. First, we should pivot it longer.

```{r}
stringency_index_long <- pivot_longer(stringency_index, -c(X1, country_code, country_name),
                                      names_to = "date", 
                                      values_to = "stringency") %>% 
  mutate(date = lubridate::dmy(date))

stringency_index_long
```

Stringency index in Denmark

```{r}
stringency_index_long %>% 
  filter(country_name == "Denmark") %>% 
  ggplot(aes(date, stringency)) + 
  geom_step() + 
  scale_x_date()
```

