---
title: "HOPE questionaire"
author: "Johannes Enevoldsen"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    fig.path = "figs/",
    message = FALSE,
    warning = FALSE
)

```

> This document is generated from [README.rmd](README.rmd)

Answers from the daily questionnaires from the [HOPE project](https://hope-project.dk/#/) will be available during the datathon.
The data is aggregated to daily counts for each answers for each question. The data is provided as a .dta (STATA) file and can be read using the [{haven}](https://haven.tidyverse.org/) package.

```{r}
library(tidyverse)
library(haven)
```

```{r}
attitudes <- read_dta("distributions_attitudes.dta")
head(attitudes)
```

## The structure

 - Date is the date of the answers.
 - Columns ending with `_labels` are labelled categorical values (similar to factors) corresponding to each possible answer to a question.
 - Remaining columns (e.g. Q1_1) counts the number of participants who have given a specific answer (corresponding the to `_labels` column left of it) on a given date.
 
 
## Label columns

The labelled columns can be converted to a factor with `labelled::to_factor`

```{r}
attitudes$attitude_labels %>% head(10)
```

```{r}
labelled::to_factor(attitudes$attitude_labels) %>% head(10)
```

Here, we make a second version of all `_labels` columns, converted to a factor:

```{r}
attitudes <- attitudes %>% 
    mutate(across(ends_with("_labels"), labelled::to_factor, levels = "labels", decreasing = TRUE, .names = "{.col}_2"))

attitudes$mask_labels_2 %>% head(10)
```

*In the first few months, there was no questions regarding mask usage*

## Calculate proportions 

The number of people who answer each day differs, we will therefore calculate the daily proportion of each answer:

```{r}
attitudes_prop <- attitudes %>% 
    group_by(date) %>% 
    mutate(across(c(starts_with("Q"), "mask", "contact_w_infected", "test", "selfisolation", "selfisolation1", "vaccination"), 
                  .fns = ~.x / sum(.x), # Since the data frame has been grouped by day, the sum will be calculated for each day.
                  ))

```

We can illustrate the daily distribution in answers using a stacked bar plot of the proportions.

```{r fig.height=4, fig.width=10}
ggplot(attitudes_prop, aes(date, Q1_1 * 100, fill = fct_relevel(attitude_labels_2, "Don't know", after = Inf))) +
    geom_col(position = "stack", width = 1) +
    labs(title = "Q1.1 - I hvilken grad føler du, at du er udsat i forhold til coronavirussen?",
         fill = "Answer",
         y = "Proportion [%]")
```

