---
title: "Mobility of mobile phones in Denmark"
author: "Johannes Enevoldsen"
output: 
    github_document:
        df_print: kable
---

    ```{r setup, include=FALSE}
knitr::opts_chunk$set(
    fig.path = "figs/",
    message = FALSE,
    warning = FALSE
)

```

> This document is generated from `README.rmd`

Mobility data based on the movement of mobile phones between telephone towers in Denmark is available from <https://covid19.compute.dtu.dk/>. The data is collected and aggregated from different network providers as described [here](https://covid19.compute.dtu.dk/data-description/telco_data/).

A data file containing the number of movements between each combination of municipalities each day can be downloaded from the website:

```{r message=FALSE}
telco_link <- "https://covid19.compute.dtu.dk/data/telco_map_new.json"

# The file is large (>17 MB) so only download if missing (or outdated)
if(!file.exists("telco_map_new.json")) {
    download.file(telco_link, destfile = "telco_map_new.json")
}
```




The data is provided as a .json file (a text based data format commonly used in web applications). We can read it into R as a list using the {jsonlite} package. We'll also use {tidyverse}.

```{r, message=FALSE}
library(jsonlite)
library(tidyverse)

telco_list <- fromJSON("telco_map_new.json") # Load json file

str(telco_list, list.len = 2) # Print structure
```
```{r, include = FALSE}
theme_set(jenevoldsen::theme_bw2())
```

The structure is a list of municipalities, that each contain a list of municipalities, that each contain a vector of number of daily trips between the pair of municipalities (the direction is unknown). `telco_list$Aalborg$Frederikshavn` are trips between Aalborg and Frederikshavn, `telco_list$Aalborg$Aalborg` are trips within Aalborg, and ``telco_list$Aalborg$`_Aalborg` `` are all trips involving Aalborg. 

The vector of daily trips (e.g. `telco_list$Aabenraa$Aalborg`) corresponds to the dates in ``telco_list$`_meta`$datetime``. I cases where the vector of daily trips is shorter than the number of days, the trips correspond to the first dates, and the missing values are zero.

We'll make these zeros explicit. Also the total entry (`$_municipality`) is a summary (sum) of the daily trips and can be calculated if needed. We'll remove that too.

```{r}
telco_list_trips <- telco_list[1:98] # skip the meta item

telco_list_trips <- map(telco_list_trips, function(x) {
    # In each sub-list, remove the list which name starts with "_"
    x[str_detect(names(x), "^_")] <- NULL
    x
    })

telco_dates <- as.Date(telco_list$`_meta`$datetime)

zero_pad_end <- function(vec, total_len) {
    stopifnot(total_len >= length(vec))
    c(vec, rep(0, total_len - length(vec)))
}

telco_list_trips <- rapply(telco_list_trips, zero_pad_end, total_len = length(telco_dates), how = "replace")
```

This data would fit nicely in a 3D array (Municipality1 x Municipality2 x date)... But this is R, so we'll jam it into a dataframe.

```{r}
telco_trips_df <- tibble(municip1 = names(telco_list_trips), 
                         municip2 = telco_list_trips) %>% 
    unnest_longer(municip2, indices_to = "municip2", values_to = "trips") %>% 
    mutate(date = list(telco_dates), .before = "municip1") %>% 
    unnest(c(trips, date)) %>% 
    relocate(municip2, .after = municip1)

head(telco_trips_df)
```

We can now easily filter and plot the trips

```{r}

filter(telco_trips_df, municip1 == "Århus", municip2 %in% c("Silkeborg", "Vejle", "Syddjurs", "Aabenraa")) %>% 
ggplot(aes(date, trips, color = municip2)) + 
    geom_line() +
    labs(title = "Daily trips to/from Aarhus")
```

We should be cautious with long trips (e.g Aarhus <--> Aabenraa), since they can end up being registered as multiple smaller trips instead. Telecompanies register end-of-trip differently, so a change in long trips may be caused by a change in reporting rather than a change in traveling behavior (e.g. in February 2021).

```{r}
filter(telco_trips_df, municip1 == "Århus", municip2 == "Aabenraa") %>% 
ggplot(aes(date, trips, color = municip2)) + 
    geom_line() +
    labs(title = "Daily trips between Aarhus and Aabenraa")
```

